---
title: "Week 6 Discussion"
author: "Michelle Vered"
date: "`r Sys.Date()`"
output: html_document
---

# Code

Initial set-up:

```{r, echo = FALSE}

# set up options & libraries
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(skimr)
library(moderndive)

# read in data 
data <- read_delim('o-ring-erosion-only.txt')
```

Exploratory Data Analysis

```{r}

data %>%
  skim()
```

# Analysis

## Links

Github Link:

Data Source Link: <https://archive.ics.uci.edu/ml/datasets/Challenger+USA+Space+Shuttle+O-Ring>

## Background on Data Set

This data set includes information related to the explosion of the Space Shuttle Challenger on January 28th, 1986. An investigation revealed that the explosion occurred because O-rings used in rocket construction failed on the day of launch. These O-rings were used to seal the joints on the booster rockets. The investigation found that the O-rings were vulnerable to failure when exposed to cold temperatures.

## Research Question & Motivation

## Data Description & Descriptive Statistics

The data set includes information

## Data Analysis & Visualizations

```{r}


```

## Findings
