---
title: "discussion1"
author: "Michelle Vered"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

# Code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

```{r}

# read in data
shoes <- read.csv('shoes.csv', na.strings=c(''))

# convert date columns to date format
shoes$dateAdded <- as.POSIXlt(shoes$dateAdded, 
                              tz = "UTC", 
                              tryFormats = c("%Y-%m-%dT%H:%M:%SZ"), 
                              optional = FALSE)
shoes$dateUpdated <- as.POSIXlt(shoes$dateUpdated, 
                              tz = "UTC", 
                              tryFormats = c("%Y-%m-%dT%H:%M:%SZ"), 
                              optional = FALSE)

# **TECHNIQUE #1 FROM THE WEEK'S MATERIAL: summarize data**
summary(shoes)

# **TECHNIQUE #2 FROM THE WEEK'S MATERIAL: count up the number of shoes per brand**
# then save table as a dataframe
brand_table <- table(shoes$brand) %>%
  as.data.frame()

# count number of NAs in each column
sapply(shoes, function(x) sum(is.na(x)))
```

```{r}
# generate a nice looking table with a list of all brands that have at least 100 pairs of shoes in the data set
brand_table %>%
  filter(Freq >= 100) %>%
  kable()

```


# Analysis

## Links

Github Link: 
Data Source Link: https://www.kaggle.com/datasets/datafiniti/womens-shoes-prices

## Description of Data

This data set contains information on women's shoes, including their brand, manufacturer, color, and price. The data was scrapped from online shopping platforms by the company Datafiniti. It was originally uploaded in 2019 based on data collected from online shopping platforms between 2015 - 2018. There are 10,000 observations in this data set, each representing a particular shoe brand, model, and retailer combination. 

Here are some key columns from the data set:
- id: A unique id for each observation. No missing values.
- dateAdded: Date the listing was first added to the data set. No missing values. The earliest date is 2015-05-04 and the most recent date is 2018-05-21. The median date was 2017-11-15.
- dateUpdated: Date the brand listing was most recently updated. No missing values. The earliest date is 2018-01-01 and the most recent date is 2018-10-23. The median date was 2018-02-01.
- brand: A string listing the name of the brand of the shoes. No missing values.
- colors: A string listing the colors available for each pair of shoes. Only about a quarter of rows have color information listed. 7,369 rows are missing color information.
- manufacturer: This is the manufacturer of the shoe, if different from the brand. Most rows (9,473) have this field empty, indicating the brand and manufacturer are the same. 
- name: A string listing the name of the shoe model. No missing values.
- prices.amountMax: The highest price listed online for this brand and model of shoe. No missing data. The the values in this column range from 5.87 - 359.99, with a mean of 69.22. 
- prices.amountMin: The lowest price listed online for this brand and model of shoe. No missing data. The values in this column range from 4.88 - 359.99, with a mean of 51.13. 
- prices.isSale: This column indicates whether an item was on sale at the time its price was recorded. This data is present for all rows in the data set. 


## Research Question & Motivation

We're interested in assessing what the average price of women's shoes are, how much the price of a pair of shoes varies between its highest and lowest listed prices, and whether shoes are actually cheaper on sale. It is possible that expensive shoes may be the ones that tend to go on sale more often, meaning shoes on sale could be more expensive than shoes not on sale. Or, conventional wisdom could hold true and sale shoes could be cheaper.

These questions would be useful for a potential purchaser of women's shoes. But this information could also be helpful if you were working for a shoe business - knowing information abut how competitors price their shoes could help you to price your shoes competitively with the marketplace.


## Findings

1. 
2.
3.



