---
title: "discussion3"
author: "Michelle Vered"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

# Code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

```{r}
# read in data
cereal <- read.csv('cereal.csv')

# TECHNIQUE #1:recode -1 values to NA
cereal <- cereal %>%
  mutate(carbo = ifelse(carbo < 0, NA, carbo)) %>%
  mutate(sugars = ifelse(sugars < 0, NA, sugars)) %>%
  mutate(potass = ifelse(potass < 0, NA, potass))

# summarize data
summary(cereal)
```

# Analysis

## Links

Github Link: <https://github.com/mvered/DMA3300/tree/main/week_3>

Data Source Link: <https://www.kaggle.com/datasets/crawford/80-cereals>

## Background on Data Set

The data includes information on breakfast cereals, including the name, manufacturer, and nutrition facts. The data was cleaned by Petra Isenberg, Pierre Dragicevic and Yvonne Jansen. The original source of the data they used was from the StatLib 1993 Statistical Graphics Exposition, where the Cereal data set was one of the data sets available to participants to use in creating data visualizations for the conference.

## Research Question & Motivation

It's useful to understand the nutrition facts of different cereals, both as a consumer of cereals and from the perspective of a cereal seller or manufacturer. From the perspective of a cereal consumer, you might be interested in understanding what kind of nutrition cereal provides - how does the sugar content in cereals correlate to the calorie count? From a seller or marketplace perspective, are there certain cereal manufacturers that have different niches within the market? For example, do some producers tend towards creating sugary cereals while others mostly produce lower-sugar cereals? Or do most brands tend to produce a mix of different types of cereals?

## Data Description & Descriptive Statistics

The data includes 77 observations, each one representing a different cereal brand and name. None of the rows or columns have missing data explicitly, but a few columns have negative values, which may indicate that the data is actually missing but coded as -1.

The data set includes the following columns:

-   name: The name of the cereal.

-   mfr: The manufacturer of the cereal, coded by letter.

    -   A = American Home Food Products

    -   G = General Mills

    -   K = Kelloggs

    -   N = Nabisco

    -   P = Post

    -   Q = Quaker Oats

    -   R = Ralston Purina

-   type: Whether the cereal is hot or cold.

-   calories: The number of calories per serving. There is an average of 106.9 calories per serving, with a minimum of 50 and a maximum of 160 calories.

-   protein: The protein per serving, in grams. There is an average of 2.545 grams, with a range of 1 to 6 grams.

-   fat: The fat per serving, in grams. There is an average of 1.013 grams, with a range of 0 to 5 grams.

-   sodium: Sodium per serving, in milligrams. There is an average of 159.7 milligrams, with a range of 0 to 320 milligrams.

-   fiber: Dietary fiber per serving, in grams. There is an average of 2.152 grams, with a range of 0 to 14 grams.

-   carbo: Complex carbohydrates per serving, in grams. **There are some rows with a value of -1, which may indicate missing data, since negative values for nutrition facts do not make sense.** If we drop the -1 values, the lowest value is 5 grams and the highest value is 23 grams. After dropping the one row coded as -1, the average value is 14.6 grams.

-   sugars: Sugars per serving, in grams. This column also includes some values of -1 which may indicate missing data. After dropping the one row coded as -1, the average value is 6.922 grams, with a range of 0 to 15 grams.

-   potass: Potassium per serving, in milligrams. This column also includes some -1 values which may indicate missing data. Dropping the two rows with -1 as their value, the average is 96.08 milligrams, with a range of 15 to 330 milligrams of potassium.

-   vitamins: The percentage of the FDA recommended daily intake of vitamins and minerals that the cereal holds, per serving. Either 0, 25, or 100. The average was 28.25.

-   shelf: The display shelf the cereal is placed on, counting from the floor up. The average was shelf number 2.208.

-   weight: The weight of one serving, in ounces. The average was 1.03 oz, with a range of 0.5 to 1.5 oz.

-   cups: The number of cups in one serving. The average was 0.821 cups, with a range of 0.25 to 1.5 cups.

-   rating: A rating of the cereal, on a scale of 0-100, from Consumer Reports. The average rating was 42.67 with a range of 18.04 to 93.70 points.

## Data Analysis & Visualizations

There is a positive correlation between sugar and calories, as we can see from the scatterplot of those two variables. The higher the sugar content, the larger the number of calories in each serving, as we might expect. The cereals tend to exist on a continous spectrum of lower to higher sugar content and lower to higher calorie content, rather than being separated into distinct \"low\" and \"high\" groups. From this visualization, it\'s hard to make out any distinct patterns by manufacturer. Each manufacturer has cereals across the spectrum of sugar content and calorie content.

```{r, echo = FALSE}

# visualization of sugar vs calorie count
cereal %>%
  drop_na() %>%
  ggplot(aes(x = sugars, y = calories, color = mfr)) +
  geom_point() +
  labs(x = "Sugar per serving, in grams",
       y = "Calories per serving",
       title = "Sugar versus Calories, by Manufacturer",
       color = "Manufacturer")
```

We can also look at summary statistics for calorie content and sugar content, grouping cereals by manufacturer. Specifically, we'll look at the mean and standard deviation of calories and sugar.

```{r, echo = FALSE}

# TECHNIQUE #2: group by & summarize
cereal %>%
  drop_na() %>%
  group_by(mfr) %>%
  summarize(count_cereals = n(),
            calories_mean = round(mean(calories),2),
            calories_sd = round(sd(calories),2),
            sugars_mean = round(mean(sugars),2),
            sugars_sd = round(sd(sugars),2)
            ) %>%
  kbl() %>% # format table nicely
  kable_classic(full_width = F) # choose style of table formatting
```

Before diving into caloric and sugar content, however, it's important to note that the number of cereals produced by a given manufacturer varies wildly. Manufacturer A (American Home Food Products) only produces 1 cereal, while at the upper end, Manufacturer K (Kellogs) produces 23 cereals. So some manufacturers have the opportunity for more variety just by virtue of having so many products, while other manufacturers are focusing on only a handful of product offerings.

In terms of calories, the average caloric content for one serving of cereal is between 84.00 to 115.71 calories per serving. The standard deviation of calories, a measure of the variation from cereal to cereal, varies depending on the manufacturer. Some manufacturers produce cereals that are more similar to one another in caloric content, such as Manufacturer N (Nabisco) which has a standard deviation of 8.904, Manufacturer G (General Mills) with a standard deviation of 10.37 and Manufacturer P (Post) with a standard deviation of 10.54. The other manufacturers have a much greater variation in caloric content between their different cereals, with a standard deviation ranging from 22.22 to 31.55.

In terms of sugar, the mean sugar content ranges from 2.20 to 8.78 grams per serving. The standard deviation of sugar content ranges from 3.03 to 4.78 grams. We can visualize the distribution of sugar content by manufacturer using a boxplot. From the boxplot, it's easy to see that Manufacturer N (Nabisco) has cereals that are far less sugary than their competitors. The other manufacturers are more similar to one another in that they have a wider range of values for sugar content. This indicates that these manufacturers are producing a variety of products, some with higher sugar content and some that are lower in sugar.

```{r, echo = FALSE}

cereal %>%
  drop_na() %>%
  ggplot(aes(x = mfr, y = sugars)) +
  geom_boxplot() +
  labs(x = "Manufacturer Code",
       y = "Sugar per Serving, in Grams",
       title = "Boxplot of Cereal Sugar Content by Manufacturer")
```

## Findings

1.  Higher amounts of sugar in a cereal is associated with a higher calorie content. We can see this from the scatterplot of sugar content vs calories per serving.
2.  Some manufacturers make cereals that are similar to one another in terms of calorie count per serving, while other manufacturers make more of a variety of cereals. We can see this by looking at the standard deviation of sugar content, grouped by manufacturer.
3.  One manufacturer, Nabisco, makes most of its cereals at a lower sugar content. One manufacturer, American Home Food Products, makes only 1 cereal, so it's hard to draw conclusions about their decisions around sugar content of their products. The rest of the manufacturers produce a variety of cereals, some lower and some higher in sugar content. We can see this through the boxplot of sugar content by manufacturer.
